Pre- and Post-Debate Democratic Primary Data: Twitter, Google Trends, and Polls

### 2019-12-11 Paper/paper-draft-Padgett-Thiam.Rmd
additional graphs

```{r, echo = FALSE, include=FALSE}
library(knitr)
library(tidyverse)
library(gtrendsR)
library(dplyr)
library(ggplot2)
library(ggmap)
library(readxl)
library(shiny)
library(rtweet)
library(tidytext)
library(stringr)
library(lubridate)
library(quanteda)
library(SentimentAnalysis)
library(GGally)
library(repmis)
library(httr)
```

@@

```{r, tidy=TRUE, warning=FALSE, message=FALSE}
#clean data
gtrends <-
  interest_over_time_all %>%
  mutate(Candidate=keyword, Data="GTrends") %>%
  group_by(Candidate, Pre_post, Data) %>%
  summarise(Percentage=mean(hits))

poll <-
  polling_data %>%
  select(Candidate, Pre_post, Percentage) %>%
  mutate(Data="Polling") %>%
  group_by(Candidate, Pre_post, Data)

@@

twitter <- as_tibble(twitter_data)
twitter <-
  twitter %>%
  group_by(Candidate, Pre_post, Data) %>%
  mutate(Percentage=Percentage*100)

@@

Then, we created a shiny app, which can be viewed by running the following code:
```{r, eval=FALSE, tidy=TRUE}
# Define UI
ui <- fluidPage(

@@

# Run the application 
shinyApp(ui = ui, server = server)

@@

When looking at the pre- and post-debate data from each source side-by-side, we can see that, while candidates who are polling higher have higher numbers of tweets and Google searches (for the most part), there is no clear relationship between the post-debate changes in polling numbers and the post-debate changes in Twitter and Google Trends data. Regardless of polling numbers, all candidates had higher Twitter positivity ratings and higher numbers of Google searches post-debate than pre-debate.

@@

It would be advantageous to survey methodologists to be able to use sources such as Twitter to replace survey data. However, we need to aknowledge the limitations preventing us from doing so. We all will agree that Twitter users do not adequately represent the population consequently introducing coverage error. Validity and measurement error just to name a few limitations we encountered with Twitter data. In lieu of replacing survey data, Twitter data can/ should be use as a supplemental data. It is important to note that we collected data for a short period of time. We believed that Twitter data can be much more useful if we had a longer timeframe. With that we would be able to perform more in depht analysis with time a key variable or even fit models. For instance, the next democratic debate is scheduled for December 19th, it would be intersting to see how the sentiment differ from the November 21st debate. On that point, candidates could use Twitter data for many other purposes such as percpetion in the Twitter world or change over time in sentiment but definitely not for election outcomes.

>>>>

```{r, echo = FALSE, include=FALSE}
library(knitr)
library(tidyverse)
library(gtrendsR)
library(dplyr)
library(ggplot2)
library(ggmap)
library(readxl)
library(shiny)
library(rtweet)
library(tidytext)
library(stringr)
library(lubridate)
library(quanteda)
library(SentimentAnalysis)
library(GGally)
library(repmis)
library(httr)
library(rsconnect)
```

@@

```{r, tidy=TRUE, warning=FALSE, message=FALSE}
#clean data
gtrends <-
  interest_over_time_all %>%
  mutate(Candidate=keyword, Data="GTrends") %>%
  filter(Candidate %in% c("Joe Biden", "Bernie Sanders", 
                          "Elizabeth Warren", "Kamala Harris", 
                          "Pete Buttigieg")) %>%
  group_by(Candidate, Pre_post, Data) %>%
  arrange(Candidate, Pre_post) %>%
  summarise(Percentage=mean(hits))

poll <-
  polling_data %>%
  select(Candidate, Pre_post, Percentage) %>%
  filter(Candidate %in% c("Joe Biden", "Bernie Sanders", 
                          "Elizabeth Warren", "Kamala Harris", 
                          "Pete Buttigieg")) %>%
  mutate(Data="Polling") %>%
  arrange(Candidate, Pre_post) %>%
  group_by(Candidate, Pre_post, Data)

@@

twitter <- as_tibble(twitter_data)
twitter <-
  twitter %>%
  group_by(Candidate, Pre_post, Data) %>%
  arrange(Candidate, desc(Pre_post)) %>%
  mutate(Percentage=Percentage*100)

@@

Then, we created a shiny app, which can be viewed at https://znpadgett.shinyapps.io/shinyapp/:
```{r, eval=FALSE, tidy=TRUE}
# Define UI
surv727app <-
ui <- fluidPage(

@@

# Run the application 
shinyApp(ui = ui, server = server)
```

```{r, eval=FALSE, include=FALSE}
#upload shiny app to the web - fake secret
rsconnect::setAccountInfo(name='znpadgett',
			  token='DA192A4A3FA4FB169FE8B04D20CA9CEA',
			  secret='XXXXXXXXXXXXXXXXXXXXXXXXXX')
```

Finally, we examined the relationships between the polling data and the Twitter and Google Trends data post-debate, as well as the relationships between pre- and post-debate changes across data sources.

```{r, tidy=TRUE, fig.cap="Comparison of Post-debate Twitter and Polling Data"}
graphing_data <- cbind(twitter, poll, gtrends)
graphing_data %>%
  group_by(Pre_post, Candidate) %>%
  filter(Pre_post=="Post-debate") %>%
  ggplot() +
  geom_point(mapping = aes(x=Percentage, y=Percentage1, color=Candidate), size=4) +
  xlim(0,20) +
  ylim(0,30) +
  labs(x="Twitter Sentiment", 
       y="Polling Data")
```

```{r, tidy=TRUE, fig.cap="Comparison of Post-debate Google Trends and Polling Data"}
graphing_data  %>%
  group_by(Pre_post, Candidate) %>%
  filter(Pre_post=="Post-debate") %>%
  ggplot() +
  geom_point(mapping = aes(x=Percentage2, y=Percentage1, color=Candidate), size=4) +
  xlim(0,100) +
  ylim(0,30) +
  labs(x="Google Trends Data", 
       y="Polling Data")
```

```{r, tidy=TRUE, fig.cap="Comparison of Changes in Twitter and Polling Data"}
print(graphing_data)
pre <-
graphing_data  %>%
  select(Pre_post, Candidate, Percentage, Percentage1, Percentage2) %>%
  group_by(Pre_post, Candidate) %>%
  filter(Pre_post=="Pre-debate") %>%
  mutate(Pre_twitter=Percentage, Pre_polling=Percentage1, Pre_gtrends=Percentage2)

post <-
  graphing_data  %>%
  select(Pre_post, Candidate, Percentage, Percentage1, Percentage2) %>%
  group_by(Pre_post, Candidate) %>%
  filter(Pre_post=="Post-debate") %>%
  mutate(Post_twitter=Percentage, Post_polling=Percentage1, Post_gtrends=Percentage2)

change_data <- cbind(pre, post)

change_data %>%
  group_by(Pre_post, Candidate) %>%
  mutate(change_twitter = ((Post_twitter - Pre_twitter) / Pre_twitter) * 100, 
         change_polling = ((Post_polling - Pre_polling) / Pre_polling) * 100, 
         change_gtrends = ((Post_gtrends - Pre_gtrends) / Pre_gtrends) * 100) %>%
  ggplot() +
  geom_point(mapping = aes(x=change_twitter, y=change_polling, color=Candidate), size=4) +
  labs(x="Percent Change in Twitter Sentiment", 
       y="Percent Change in Polling Data")

@@

```{r, tidy=TRUE, fig.cap="Comparison of Changes in Google Trends and Polling Data"}
change_data %>%
  group_by(Pre_post, Candidate) %>%
  mutate(change_twitter = ((Post_twitter - Pre_twitter) / Pre_twitter) * 100, 
         change_polling = ((Post_polling - Pre_polling) / Pre_polling) * 100, 
         change_gtrends = ((Post_gtrends - Pre_gtrends) / Pre_gtrends) * 100) %>%
  ggplot() +
  geom_point(mapping = aes(x=change_gtrends, y=change_polling, color=Candidate), size=4) +
  labs(x="Percent Change in Google Trends Data", 
       y="Percent Change in Polling Data")
```
When looking at the pre- and post-debate data from each source side-by-side, we can see that, while candidates who are polling higher have higher numbers of tweets and Google searches (for the most part), there is no clear relationship between the post-debate changes in polling numbers and the post-debate changes in Twitter and Google Trends data. Regardless of polling numbers, all candidates had higher Twitter positivity ratings and higher numbers of Google searches post-debate than pre-debate.

@@

It would be advantageous to survey methodologists to be able to use sources such as Twitter and Google Trends to replace survey data. However, we need to acknowledge the limitations preventing us from doing so. We all will agree that Twitter users do not adequately represent the population consequently introducing coverage error. Validity and measurement error just to name a few limitations we encountered with Twitter data. In lieu of replacing survey data, Twitter data can and should be use as a supplemental data. It is also important to note that we collected data for a short period of time. We believed that Twitter data can be much more useful if we had a longer timeframe. With that we would be able to perform more in depht analysis with time a key variable or even fit models. For instance, the next democratic debate is scheduled for December 19th, it would be intersting to see how the sentiment differ from the November 21st debate. On that point, candidates could use Twitter data for many other purposes such as percpetion in the Twitter world or change over time in sentiment but definitely not for election outcomes.
